Tropical Andes IUCN Montane and Lowland forest

### 2023-07-30 code/L0/L0_TropicalAndes_IUCNhabitatClassification.Rmd
add clip extent and masking habitat types not forest

@@

# Load required packages
```{r}
library(raster)
```

# Load in Ecosystem Functional Groups
Data downloaded from www.zendo.org/record/5090419
```{r}
iucn_habitats <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004.tif"))
```

>>>>

---
title: "Tropical Andes IUCN Montane and Lowland forest"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Create raster of IUCN habitat classification Forest-Subtropical moist montane and Forest - Subtropical moist lowland for Tropical Andes."
data input: "iucn_habitatclassification_composite_lvl2_ver004.tif"
data output: "TropicalAndes_IUCNHabitat_Forest.GTiff"
date: "2023-07-18"
output: html_document
---

@@

# Load required packages
```{r}
library(raster)
library(sf)
library(rnaturalearth)
library(dplyr)
```

# Load in Ecosystem Functional Groups
Data downloaded from https://zenodo.org/record/4058819
Citation: Martin Jung, Prabhat Raj Dahal, Stuart H. M. Butchart, Paul F. Donald, Xavier De Lamo, Myroslava Lesiv, Valerie Kapos, Carlo Rondinini, & Piero Visconti. (2020). A global map of terrestrial habitat types (Version 004) [Data set]. Zenodo. https://doi.org/10.5281/zenodo.4058819

```{r}
iucn_habitats <- raster(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/iucn_habitatclassification_composite_lvl2_ver004.tif"))
iucn_habitats_colors <- read.csv(file.path(data_path, "iucn_habitatclassification_composite_lvl2_ver004/color_classification.csv"))
```

```{r}
plot(iucn_habitats)
```

Crop extent - trim to Andean States
```{r}
# Pull in world map
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")

# Subset world map. In this case we are removing the Galapagos by defining the bounding box around the Ecuador polygon.
study_region <-  worldMap %>% filter(sovereignt == "Bolivia" | sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
plot(study_region$geometry)
```

```{r}
iucn_habitats_crop <- crop(iucn_habitats, study_region)
TA_iucn_habitats <- mask(iucn_habitats_crop, study_region)
plot(TA_iucn_habitats)
```

value 109 is Forest-Subtropical moist montane & value 106 is Forest - Subtropical moist lowland
see level2.clr or lvl2_style.qml file
```{r}
#rename for masking
TA_iucn_habitats_mask <- TA_iucn_habitats
TA_iucn_habitats_mask[TA_iucn_habitats$iucn_habitatclassification_composite_lvl2_ver004 != 106 | TA_iucn_habitats$iucn_habitatclassification_composite_lvl2_ver004 != 109] <- NA
TA_iucn_habitats_forest <- mask(TA_iucn_habitats, TA_iucn_habitats_mask)

plot(TA_iucn_habitats_forest)
```

# Save raster
```{r}
TA_icunHabitat_path <- file.path(output_path, "TropicalAndes_IUCNHabitat_Forest.GTiff")
writeRaster(TA_iucnHabitats, filename = TA_icunHabitat_path, format = "GTiff", overwrite=TRUE)
```
